---
title: "Part 2 Basic Inferential Data_Analysis"
output: pdf_document
author: Volha Leusha
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

In this paper analysis of the ToothGrowth dataset from Rdataset package is presented and basic summary of data is provided. In second part confidence intervals and hypothesis tests are used to compare tooth growth by supp and dose, with subsequent conclusions on results.


##Dataset Base Analysis

1. Load dataset from datasets library:  

```{r}
#Load the ToothGrowth data: 
library(datasets)
data("ToothGrowth")
```

2. Basic summary of data exploratory data analysis:
```{r}
#first glance on the data:
head(ToothGrowth)

#look into data description:
str(ToothGrowth)
summary(ToothGrowth)
```

The dataset is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

Dataset has 60 observarions of 3 variables:  

- supp: factor with only two possible supplement types: OJ, VC with equal proportion
- dose: numerical categorical variable that can take only 3 unique values (milligrams/day)
```{r}
#check for unique dose values
unique(ToothGrowth$dose)
```
- len: numerical tooth length in range from 4.2 to 33.90 with mean 18.81 and standard deviation
```{r}
#find standard deviation for length
sd(ToothGrowth$len)
```
len variable has following density distribution:

```{r}
#open ggplot2 and gripExtra libraries:
library(ggplot2)
library(gridExtra)

#Plot general tooth length:
g1<- ggplot(data=ToothGrowth, aes(ToothGrowth$len))+geom_density()+
    labs(title="General toothlength", x="")

#Plot tooth length by supplements
g2<- ggplot(data=ToothGrowth, aes(ToothGrowth$len, color=ToothGrowth$supp))+
    geom_density()+labs(title="Tooth length by supplements ", x="")

#Plot tooth length by dosage
g3<- ggplot(data=ToothGrowth, aes(ToothGrowth$len, color =as.character(ToothGrowth$dose)))+
    geom_density()+labs(title="Tooth length by dosage", x="")

#arrange all three graphs on one page 
grid.arrange(g1, g2, g3, nrow = 3)
```

From graphs can be seen that data is skewed in general for all groups.

## Tooth growth comparisson by supp and dose

Plot dataset grouped by supplement type and dosage:

```{r}
#Plot dataset grouped by supplement type and dosage
ggplot(ToothGrowth, aes(x=dose, y=len, shape=supp, color=supp)) +geom_point()+
    theme(plot.margin = margin(3,3,3,3, "cm"))
```

From the graph and table it can be seen that there is potential difference in tooth length between supplements, suggesting that pigs on supplement OJ  have longer teeth than on VC. Also, the bigger the dossage, the longer are teeth. To prove above assumptions, three t-tests are performed.

1. First test:  
Ho: there is no difference between tooth length with different supplements 
Ha: tooth length for supplement OJ in higher  
alpha =.05  

```{r}
#create new variable diff1 equal to difference between values of length with OJ 
#and VC supplements (each supplement has equal amount of values in dataset= 30):
diff1<-ToothGrowth$len[ToothGrowth$supp=="OJ"]-ToothGrowth$len[ToothGrowth$supp=="VC"]
t.test(diff1, data= ToothGrowth)$conf
```
Confidence interval is fully above zero (and p-value is less than alpha)  consequently null hypothesis can be rejected and it can be assumed that pig with supplement OJ have bigger teeth length.

2. Second test:  
Ho: there is no difference between dose0.5 and 1.0 for tooth length
Ha: tooth length for dose 1.0 is higher than for dose 0.5  
alpha= .05  

```{r}
#create new variable diff2 equal to difference between values of dose 1.0
# and dose 0.5 (each dose has equal amount of values in dataset = 20):
diff2<- ToothGrowth$len[ToothGrowth$dose==1]-ToothGrowth$len[ToothGrowth$dose==.5]
t.test(diff2, data= ToothGrowth)
```

Confidence interval is fully above zero (and p-value is less than alpha)  consequently null hypothesis can be rejected and it can be assumed that Ha is true.

3. Third test:  
Ho: there is no difference between dose 1.0 and 2.0 for tooth length
Ha: tooth length for dose 2.0 is higher than for dose 1.0  
alpha= .05  

```{r}
#create new variable diff2 equal to difference between values of dose 2.0
# and dose 1.0 (each dose has equal amount of values in dataset = 20):
diff3<- ToothGrowth$len[ToothGrowth$dose==2]-ToothGrowth$len[ToothGrowth$dose==1]
t.test(diff3, data= ToothGrowth)
```

Confidence interval is fully above zero (and p-value is less than alpha)  consequently null hypothesis can be rejected and it can be assumed that Ha is true.
